On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-07-26 Manuscript/manuscript.tex
Responded to some comments

%Abstract should be 200 words
%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
%models, venue effect, and conclusion
\begin{abstract}
Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
World Track and Field Championships after registering a reaction time of 0.099 
seconds, 0.001 seconds faster than what is allowed. Following the games, 
bloggers on the running website LetsRun concluded that the reaction time data 
from the 2022 World Championships seemed to be generally faster compared to the 
other datasets they considered, but they did not perform any formal 
statistical analyses. This paper questions the reaction time 
disqualification barrier, which is currently 0.1 seconds, to determine whether 
this is a reasonable threshold. We employ a generalized linear mixed model 
(GLMM) with a random venue effect in order to  model the reaction time data. 
Additionally, we employ a signed-rank test for clustered data to 
compare reaction times for the same athletes at different competitions. This 
matter needs to be addressed because disqualification based on allowable 
reaction time will continue to be an issue in future world championships.

@@

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with a random effect 
accounting for the venue (or year) and a second random effect accounting for the
heat effect within each year. \jy{do a thorough check to make sure this is updated.}
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 reaction times differed significantly from 
those of other competitions.
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than 0.1, 0.08, and
  0.08 for various data sets.}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.08) \\ 
   \midrule
   Old Finals & $4.36\cdot10^{-4}$ & $3.0\cdot10^{-7}$ \\
   All Finals & $8.64\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
   Old Pooled & $4.15\cdot10^{-3}$ & $4.67\cdot10^{-5}$ \\ 
   All Pooled & $5.15\cdot10^{-3}$ & $6.01\cdot10^{-5}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on probability of observing
  a time less the barrier. \of{Not sure if this makes sense}}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ x = .001) & P(RxnTime $<$ x = 0.0001) \\ 
   \midrule
   Old Finals & $0.103$ & $0.095$ \\
   All Finals & $0.101$ & $0.093$ \\
   Old Pooled & $0.093$ & $0.083$ \\ 
   All Pooled & $0.091$ & $0.082$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}

\begin{table}
  \centering
  \caption{Compares the various effect models by dataset, all probabilities are
  for observing a reaction time less than 0.1 seconds. In general the model with
  the highest probability (least extreme reaction time) is the Year Effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & Year Effect & Batch Effect & Both Effects \\ 
   \midrule
   Old Finals & $4.39\cdot10^{-4}$ & $4.24\cdot10^{-4}$ & $4.36\cdot10^{-4}$ \\
   All Finals & $8.63\cdot10^{-4}$ & $8.65\cdot10^{-4}$ & $8.64\cdot10^{-4}$ \\
   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:EffectComparison}
\end{table}

>>>>

%Abstract should be 200 words
%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
%models, venue effect, and conclusion
\begin{abstract}
Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
World Track and Field Championships after registering a reaction time of 0.099 
seconds, 0.001 seconds faster than what is allowed. Following the games, 
bloggers on the running website LetsRun concluded that the reaction time data 
from the 2022 World Championships seemed to be generally faster compared to the 
other datasets they considered, but they did not perform any formal 
statistical analysis. This paper questions the reaction time 
disqualification barrier, which is currently 0.1 seconds, to determine whether 
this is a reasonable threshold. We employ a generalized linear mixed model 
(GLMM) with a random venue effect in order to  model the reaction time data. 
Additionally, we employ a signed-rank test for clustered data to 
compare reaction times for the same athletes at different competitions. This 
matter needs to be addressed because disqualification based on allowable 
reaction time will continue to be an issue in future world championships.

@@

The objective to this paper is to analyze the reaction times at the 2022 World 
Track and Field Championships to check for any abnormalities, and we approach
this in two different ways.  First we model the data, which are
the reaction times of every non-disqualification since the 1999 World 
Championships, using a gamma linear mixed-effects model with two random effects 
one effect for the year and a second effect for the
heat within each year. \jy{do a thorough check to make sure this is updated.}
\of{This should be better now.}
Based on the fitted model, we can then estimate the probability of a reaction 
time being below a certain threshold (e.g., 0.1). 
Second, we compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship.  By
comparing reaction times from the same athlete across multiple competitions we 
are able to determine if the 2022 reaction times differed significantly from 
those of other competitions.
Our investigation into reaction times is important not only because of its
implications for future races, but also because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times, which is the goal of
every sprinter who steps onto the track.

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than 0.1 and 0.08
  for various data sets.}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.08) \\ 
   \midrule
   Old Pooled & $4.15\cdot10^{-3}$ & $4.67\cdot10^{-5}$ \\ 
   All Pooled & $5.15\cdot10^{-3}$ & $6.01\cdot10^{-5}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

\begin{table}
  \centering
  \caption{Suggested reaction time barriers based on probability of observing
  a time less the barrier. \of{Not sure if this makes sense}}
  \begin{tabular}{c c c} 
   \toprule
   Data Set & P(RxnTime $<$ x = .001) & P(RxnTime $<$ x = 0.0001) \\ 
   \midrule
   Old Pooled & $0.093$ & $0.083$ \\ 
   All Pooled & $0.091$ & $0.082$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_time}
\end{table}

%\begin{table}
  %\centering
  %\caption{Compares the various effect models by dataset, all probabilities are
  %for observing a reaction time less than 0.1 seconds. In general the model with
  %the highest probability (least extreme reaction time) is the Year Effect model.}
  %\begin{tabular}{c c c c} 
  % \toprule
  % Data Set & Year Effect & Batch Effect & Both Effects \\ 
   %\midrule
   %Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
   %All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
   %\bottomrule
  %\end{tabular}
  %\label{tab:EffectComparison}
%\end{table}
